---
title: "rjags"
output: html_document
---

#Packages
```{r}
library(rjags)
library(R2jags)
library(runjags)
library(tidyverse)
library(matrixStats)
library(ROCit)
```

# Example
```{r}
model.file <- system.file(package = "R2jags", "model", "schools.txt")
J <- 8.0
y <- c(28.4,7.9,-2.8,6.8,-0.6,0.6,18.0,12.2)
sd <- c(14.9,10.2,16.3,11.0,9.4,11.4,10.4,17.6)

jags.data <- list("y","sd","J")
jags.params <- c("mu","sigma","theta")
jags.inits <- function(){list("mu"=rnorm(1),"sigma"=runif(1),"theta"=rnorm(J))}

jagsfit <- jags(data=list("y","sd","J"), inits = jags.inits,jags.params, n.iter = 10, model.file = model.file)

jagsfit
```



# BIRD model
```{r}
bird <- "model{
for(j in 1:N_VARIANTS){
  p[j] ~ dbeta(v[j]*(c2-2)+1, (1-v[j])*(c2-2)+1)
  theta[j] ~ dlnorm(0,s)
  for (i in 1:N_RNA){
    qi[j,i] ~ dbeta((theta[j]*p[j]/(1.0-p[j]+theta[j]*p[j]))*(c1-2)+1,(1-(theta[j]*p[j]/(1.0-p[j]+theta[j]*p[j])))*(c1-2)+1)
    k[j,i]  ~ dbin(qi[j,i],kk[j,i]+m[j,i])
    }
  for (i in 1:N_DNA){
    a[j,i]  ~ dbin(p[j],aa[j,i]+b[j,i])
    }
  }
  c1 ~ dgamma(1.1, 0.005)
  c2 ~ dgamma(1.1, 0.005)
  s  ~ dgamma(1.1,3)
}"

cat("model{
for(j in 1:N_VARIANTS){
  p[j] ~ dbeta(v[j]*(c2-2)+1, (1-v[j])*(c2-2)+1)
  theta[j] ~ dlnorm(0,s)
  for (i in 1:N_RNA){
    qi[j,i] ~ dbeta((theta[j]*p[j]/(1.0-p[j]+theta[j]*p[j]))*(c1-2)+1,(1-(theta[j]*p[j]/(1.0-p[j]+theta[j]*p[j])))*(c1-2)+1)
    k[j,i]  ~ dbin(qi[j,i],kk[j,i]+m[j,i])
    }
  for (i in 1:N_DNA){
    a[j,i]  ~ dbin(p[j],aa[j,i]+b[j,i])
    }
  }
  c1 ~ dgamma(1.1, 0.005)
  c2 ~ dgamma(1.1, 0.005)
  s  ~ dgamma(1.1,3)
}", file="BIRD.txt")

```

#Data
```{r}
rawdata <- read.csv("/Users/sean00002/Dropbox/PhD/Andrew_Rotation/STANINPUTS/2_data_0.01_1000_read5050.txt", sep = "\t", header = F)

bird_data_001_50 <- list(v = rawdata$V1, a = matrix(rawdata$V7),aa = matrix(rawdata$V7), b = matrix(rawdata$V6), k = cbind(rawdata$V10, rawdata$V12,rawdata$V14,rawdata$V16,rawdata$V18,rawdata$V20,rawdata$V22,rawdata$V24,rawdata$V26,rawdata$V28),kk  = cbind(rawdata$V10, rawdata$V12,rawdata$V14,rawdata$V16,rawdata$V18,rawdata$V20,rawdata$V22,rawdata$V24,rawdata$V26,rawdata$V28), m = cbind(rawdata$V9, rawdata$V11,rawdata$V13,rawdata$V15,rawdata$V17,rawdata$V19,rawdata$V21,rawdata$V23,rawdata$V25,rawdata$V27), N_VARIANTS = nrow(rawdata), N_RNA=10,N_DNA=1)

rawdata <- read.csv("/Users/sean00002/Dropbox/PhD/Andrew_Rotation/STANINPUTS/2_data_0.01_1000_read3030.txt", sep = "\t", header = F)

bird_data_001_30 <- list(v = rawdata$V1, a = matrix(rawdata$V7),aa = matrix(rawdata$V7), b = matrix(rawdata$V6), k = cbind(rawdata$V10, rawdata$V12,rawdata$V14,rawdata$V16,rawdata$V18,rawdata$V20,rawdata$V22,rawdata$V24,rawdata$V26,rawdata$V28),kk  = cbind(rawdata$V10, rawdata$V12,rawdata$V14,rawdata$V16,rawdata$V18,rawdata$V20,rawdata$V22,rawdata$V24,rawdata$V26,rawdata$V28), m = cbind(rawdata$V9, rawdata$V11,rawdata$V13,rawdata$V15,rawdata$V17,rawdata$V19,rawdata$V21,rawdata$V23,rawdata$V25,rawdata$V27), N_VARIANTS = nrow(rawdata), N_RNA=10,N_DNA=1)

rawdata <- read.csv("/Users/sean00002/Dropbox/PhD/Andrew_Rotation/STANINPUTS/2_data_0.05_1000_read5050.txt", sep = "\t", header = F)

bird_data_005_50 <- list(v = rawdata$V1, a = matrix(rawdata$V7),aa = matrix(rawdata$V7), b = matrix(rawdata$V6), k = cbind(rawdata$V10, rawdata$V12,rawdata$V14,rawdata$V16,rawdata$V18,rawdata$V20,rawdata$V22,rawdata$V24,rawdata$V26,rawdata$V28),kk  = cbind(rawdata$V10, rawdata$V12,rawdata$V14,rawdata$V16,rawdata$V18,rawdata$V20,rawdata$V22,rawdata$V24,rawdata$V26,rawdata$V28), m = cbind(rawdata$V9, rawdata$V11,rawdata$V13,rawdata$V15,rawdata$V17,rawdata$V19,rawdata$V21,rawdata$V23,rawdata$V25,rawdata$V27), N_VARIANTS = nrow(rawdata), N_RNA=10,N_DNA=1)

rawdata <- read.csv("/Users/sean00002/Dropbox/PhD/Andrew_Rotation/STANINPUTS/2_data_0.05_1000_read3030.txt", sep = "\t", header = F)

bird_data_005_30 <- list(v = rawdata$V1, a = matrix(rawdata$V7),aa = matrix(rawdata$V7), b = matrix(rawdata$V6), k = cbind(rawdata$V10, rawdata$V12,rawdata$V14,rawdata$V16,rawdata$V18,rawdata$V20,rawdata$V22,rawdata$V24,rawdata$V26,rawdata$V28),kk  = cbind(rawdata$V10, rawdata$V12,rawdata$V14,rawdata$V16,rawdata$V18,rawdata$V20,rawdata$V22,rawdata$V24,rawdata$V26,rawdata$V28), m = cbind(rawdata$V9, rawdata$V11,rawdata$V13,rawdata$V15,rawdata$V17,rawdata$V19,rawdata$V21,rawdata$V23,rawdata$V25,rawdata$V27), N_VARIANTS = nrow(rawdata), N_RNA=10,N_DNA=1)

rawdata <- read.csv("/Users/sean00002/Dropbox/PhD/Andrew_Rotation/STANINPUTS/2_data_1.0_1000_read3030.txt", sep = "\t", header = F)

bird_data_30 <- list(v = rawdata$V1, a = matrix(rawdata$V7),aa = matrix(rawdata$V7), b = matrix(rawdata$V6), k = cbind(rawdata$V10, rawdata$V12,rawdata$V14,rawdata$V16,rawdata$V18,rawdata$V20,rawdata$V22,rawdata$V24,rawdata$V26,rawdata$V28),kk  = cbind(rawdata$V10, rawdata$V12,rawdata$V14,rawdata$V16,rawdata$V18,rawdata$V20,rawdata$V22,rawdata$V24,rawdata$V26,rawdata$V28), m = cbind(rawdata$V9, rawdata$V11,rawdata$V13,rawdata$V15,rawdata$V17,rawdata$V19,rawdata$V21,rawdata$V23,rawdata$V25,rawdata$V27), N_VARIANTS = nrow(rawdata), N_RNA=10,N_DNA=1)


rawdata <- read.csv("/Users/sean00002/Dropbox/PhD/Andrew_Rotation/STANINPUTS/2_data_1.0_1000_read5050.txt", sep = "\t", header = F)

bird_data_50 <- list(v = rawdata$V1, a = matrix(rawdata$V7),aa = matrix(rawdata$V7), b = matrix(rawdata$V6), k = cbind(rawdata$V10, rawdata$V12,rawdata$V14,rawdata$V16,rawdata$V18,rawdata$V20,rawdata$V22,rawdata$V24,rawdata$V26,rawdata$V28),kk  = cbind(rawdata$V10, rawdata$V12,rawdata$V14,rawdata$V16,rawdata$V18,rawdata$V20,rawdata$V22,rawdata$V24,rawdata$V26,rawdata$V28), m = cbind(rawdata$V9, rawdata$V11,rawdata$V13,rawdata$V15,rawdata$V17,rawdata$V19,rawdata$V21,rawdata$V23,rawdata$V25,rawdata$V27), N_VARIANTS = nrow(rawdata), N_RNA=10,N_DNA=1)
```

# runjags
## Model_30 1000 variants with theta=2, DNA depth = RNA depth = 30
```{r}
ptm <- proc.time()
model_30 <- run.jags(
    model=bird, 
    monitor=c("theta"), 
    data=bird_data_30,n.chains = 4, burnin = 1000, sample=1000,method = 'parallel'
)
proc.time() - ptm 
```

## Model_30 Results
```{r}
mcmc <- as.mcmc(model_30)
mcmc <- as.matrix(mcmc)
medians = colMedians(mcmc)
ggplot(data = as.data.frame(medians), aes(x = medians)) + geom_histogram(aes(y = ..density..)) + geom_density() + geom_vline(xintercept = median.default(medians))  + ggtitle(paste("median = ",median.default(medians)))
```

## Model_50 1000 variants with theta=2, DNA depth = RNA depth = 50
```{r}
ptm <- proc.time()
model_50 <- run.jags(
    model=bird, 
    monitor=c("theta"), 
    data=bird_data_50,n.chains = 4, burnin = 1000, sample=1000,method = 'parallel'
)
proc.time() - ptm 
```

## Model_50 results
```{r}
mcmc <- as.mcmc(model_50)
mcmc <- as.matrix(mcmc)
medians = colMedians(mcmc)
ggplot(data = as.data.frame(medians), aes(x = medians)) + geom_histogram(aes(y = ..density..)) + geom_density() + geom_vline(xintercept = median.default(medians))  + ggtitle(paste("median = ",median.default(medians)))
```


## model_001_50:10 variants out of 1000 with theta=2, 990 variants out of 1000 with theta=1, DNA depth = RNA depth = 50
```{r}
ptm <- proc.time()
model_001_50 <- run.jags(
    model=bird, 
    monitor=c("theta"), 
    data=bird_data_001_50,n.chains = 4, burnin = 1000, sample=1000,method = 'parallel'
)
proc.time() - ptm 
```
## Model_001_50 results 
```{r}
mcmc <- as.mcmc(model_001_50)
mcmc <- as.matrix(mcmc)
medians = colMedians(mcmc)

pos = 10
label = c(rep(1,pos), rep(0,1000-pos))

p <- c()
for (i in 1:1000){
  preg <- sum(mcmc[,i] > 1)/4000
  p <- append(p,preg)
}

ROCit_obj <- rocit(score=p,class=label,method = "empirical")
plot(ROCit_obj)
title(main = paste0("JAGS:Model_001_50; ",summary(ROCit_obj)[4,]))

```


## model_001_30  10 variants out of 1000 with theta=2, 990 variants out of 1000 with theta=1, DNA depth = RNA depth = 30
```{r}
ptm <- proc.time()
model_001_30 <- run.jags(
    model=bird, 
    monitor=c("theta"), 
    data=bird_data_001_30,n.chains = 4, burnin = 1000, sample=1000,method = 'parallel'
)
proc.time() - ptm 
```

## Model_001_30 results 
```{r}
mcmc <- as.mcmc(model_001_30)
mcmc <- as.matrix(mcmc)
medians = colMedians(mcmc)

pos = 10
label = c(rep(1,pos), rep(0,1000-pos))

p <- c()
for (i in 1:1000){
  preg <- sum(mcmc[,i] > 1)/4000
  p <- append(p,preg)
}

ROCit_obj <- rocit(score=p,class=label,method = "empirical")
plot(ROCit_obj)
title(main = paste0("JAGS:model_001_30; ",summary(ROCit_obj)[4,]))

```

## model_005_50  50 variants out of 1000 with theta=2, 950 variants out of 1000 with theta=1, DNA depth = RNA depth = 50
```{r}
ptm <- proc.time()
model_005_50 <- run.jags(
    model=bird, 
    monitor=c("theta"), 
    data=bird_data_005_50,n.chains = 4, burnin = 1000, sample=1000,method = 'parallel'
)
proc.time() - ptm 


```

## model_005_50 results 
```{r}
mcmc <- as.mcmc(model_005_50)
mcmc <- as.matrix(mcmc)
medians = colMedians(mcmc)

pos = 50
label = c(rep(1,pos), rep(0,1000-pos))

p <- c()
for (i in 1:1000){
  preg <- sum(mcmc[,i] > 1)/4000
  p <- append(p,preg)
}

ROCit_obj <- rocit(score=p,class=label,method = "empirical")
plot(ROCit_obj)
title(main = paste0("JAGS:model_005_50; ",summary(ROCit_obj)[4,]))

```


## model_005_30  50 variants out of 1000 with theta=2, 950 variants out of 1000 with theta=1, DNA depth = RNA depth = 30
```{r}
ptm <- proc.time()
model_005_30 <- run.jags(
    model=bird, 
    monitor=c("theta"), 
    data=bird_data_005_30,n.chains = 4, burnin = 1000, sample=1000,method = 'parallel'
)
proc.time() - ptm 
```

## model_005_30 results 
```{r}
mcmc <- as.mcmc(model_005_30)
mcmc <- as.matrix(mcmc)
medians = colMedians(mcmc)

pos = 50
label = c(rep(1,pos), rep(0,1000-pos))

p <- c()
for (i in 1:1000){
  preg <- sum(mcmc[,i] > 1)/4000
  p <- append(p,preg)
}

ROCit_obj <- rocit(score=p,class=label,method = "empirical")
plot(ROCit_obj) 
title(main = paste0("JAGS:model_005_30; ",summary(ROCit_obj)[4,]))
```






